---
title: "nc_tutorial"
author: "Lisa"
date: "2026-06-12"
output: html_document
---

This tutorial includes the codes developed for the main analyses used in the manuscript "Neural Idiosyncrasy Predicts Future Social Disconnectedness", run on simulated datasets with 45 dyads (10 participants) and 5 brain regions. 

The simulated intersubject similarity (ISC) data is in isc_data_simulated.csv and the simulated subject information (i.e., subject ID, indegree popularity as well as perceived popularity at Time 1 and Time 2, class affiliation, middle school affiliation, and handedness) is in subject_info_simulated.csv.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up
```{r Load packages, echo = False}
library(plyr)
library(dplyr)
library(ggplot2)
library(igraph)
library(RColorBrewer)
library(gridExtra)
library(emmeans)
library(lmerTest)
library(reshape2)
library(tidyr)
```

```{r Read in data, echo=FALSE}
isc_data <- read.csv("isc_data_simulated.csv") # may need to change path to where the data is saved
subject_info <- read.csv("subject_info_simulated.csv") # may need to change path to where the data is saved
```

```{r Recode ISC outliers}
RecodeOutliersGrouped_df <- function(df, value_col, group_col, new_col = "isc_recoded") {
  # Rename columns temporarily for processing
  temp_df <- df[, c(value_col, group_col)]
  colnames(temp_df) <- c("value", "group")
  temp_df$index <- seq_len(nrow(temp_df))
  
  # Group and recode
  recoded <- do.call(rbind, lapply(split(temp_df, temp_df$group), function(subdf) {
    q1 <- quantile(subdf$value, 0.25, na.rm = TRUE)
    q3 <- quantile(subdf$value, 0.75, na.rm = TRUE)
    iqr <- IQR(subdf$value, na.rm = TRUE)
    
    lower <- q1 - 1.5 * iqr
    upper <- q3 + 1.5 * iqr
    
    subdf$recode_value <- subdf$value
    subdf$recode_value[subdf$value < lower] <- lower
    subdf$recode_value[subdf$value > upper] <- upper
    
    return(subdf)
  }))
  
  # Restore original order
  recoded <- recoded[order(recoded$index), ]
  
  # Append new column to original df
  df[[new_col]] <- recoded$recode_value
  
  return(df)
}

isc_recoded <- RecodeOutliersGrouped_df(isc_data, 'isc', 'brain_region')
```

```{r Combine dataframes}
dyad_list <- unique(isc_recoded[,c("subject1", "subject2", "brain_region","isc_recoded")])
dyad_df <- merge(dyad_list, subject_info, by.x = c("subject1"), by.y = c("subject"))
colnames(dyad_df)[c(5:11)] <- paste("sub1", colnames(dyad_df)[c(5:11)], sep = "_")
dyad_df <- merge(dyad_df, subject_info, by.x = c("subject2"), by.y = c("subject"))
colnames(dyad_df)[c(12:18)] <- paste("sub2", colnames(dyad_df)[c(12:18)], sep = "_")

master_df <- dyad_df %>% mutate(
  dyad_indegree_min_t1 = ifelse(sub1_indegree_t1 > sub2_indegree_t1, sub2_indegree_t1, sub1_indegree_t1),
  dyad_indegree_min_t2 = ifelse(sub1_indegree_t2 > sub2_indegree_t2, sub2_indegree_t2, sub1_indegree_t2),
  dyad_perceived_min_t1 = ifelse(sub1_perceived_t1 > sub2_perceived_t1, sub2_perceived_t1, sub1_perceived_t1),
  dyad_perceived_min_t2 = ifelse(sub1_perceived_t2 > sub2_perceived_t2, sub2_perceived_t2, sub1_perceived_t2),
  same_class = as.factor(ifelse(sub1_class == sub2_class, 1, 0)),
  same_middleschool = as.factor(ifelse(sub1_middle_school == sub2_middle_school, 1, 0)),
  same_handedness = as.factor(ifelse(sub1_handedness == sub2_handedness, 1, 0))
)
```


```{r Double dataframe}
double_dataframe <- function(df_to_be_doubled){
  df_2 <- df_to_be_doubled
  colnames(df_2)[1:2] <- c("subject1", "subject2")
  df_doubled <- rbind(df_to_be_doubled, df_2)
  return(df_doubled)
}

master_df_doubled <- double_dataframe(master_df)
```

## ANALYSES
# Testing if pre-existing neural idiosyncrasy predicts lower future in-degree centrality
```{r Analysis: testing if pre-existing neural idiosyncrasy predicts lower future in-degree centrality}
corrected_df_n <- nrow(master_df)/5

func_isc_indeg_min_z <- function(data, corrected_df){
    md <- summary(lmer(scale(dyad_indegree_min_t2) ~ scale(isc_recoded) + same_class + same_middleschool + same_handedness + (1|subject1) + (1|subject2), data))

  B <- md$coefficients[[2,1]]
  #df <- md$coefficients[[2,3]]
  df <- corrected_df - 4
  p <- md$coefficients[[2,5]]
  t <- md$coefficients[[2,4]]
  p_penalized <- 2 * pt(-abs(t), df/2)
  
  return(data.frame(B, df, p, p_penalized))
}

func_isc_indeg_min_z_contr <- function(data, corrected_df){
    md <- summary(lmer(scale(dyad_indegree_min_t2) ~ scale(isc_recoded) + scale(dyad_indegree_min_t1) + same_class + same_middleschool + same_handedness + (1|subject1) + (1|subject2), data))

  B <- md$coefficients[[2,1]]
  #df <- md$coefficients[[2,3]]
  df <- corrected_df - 5
  p <- md$coefficients[[2,5]]
  t <- md$coefficients[[2,4]]
  p_penalized <- 2 * pt(-abs(t), df/2)
  
  return(data.frame(B, df, p, p_penalized))
}

results_isc_indeg_min <- ddply(master_df_doubled, .(brain_region), func_isc_indeg_min_z, corrected_df_n)
results_isc_indeg_min$corrected_p <- p.adjust(results_isc_indeg_min$p_penalized, method = "fdr")
results_isc_indeg_min_contr <- ddply(master_df_doubled, .(brain_region), func_isc_indeg_min_z_contr, corrected_df_n)
results_isc_indeg_min_contr$corrected_p <- p.adjust(results_isc_indeg_min_contr$p_penalized, method = "fdr")

results_isc_indeg_min
results_isc_indeg_min_contr
```

# Testing if pre-existing neural idiosyncrasy predicts lower future perceived popularity
```{r Analysis: testing if pre-existing neural idiosyncrasy predicts lower future perceived popularity}
corrected_df_n <- nrow(master_df)/5

func_isc_perceived_min_z <- function(data, corrected_df){
    md <- summary(lmer(scale(dyad_perceived_min_t2) ~ scale(isc_recoded) + same_class + same_middleschool + same_handedness + (1|subject1) + (1|subject2), data))

  B <- md$coefficients[[2,1]]
  #df <- md$coefficients[[2,3]]
  df <- corrected_df - 4
  p <- md$coefficients[[2,5]]
  t <- md$coefficients[[2,4]]
  p_penalized <- 2 * pt(-abs(t), df/2)
  
  return(data.frame(B, df, p, p_penalized))
}

func_isc_perceived_min_z_contr <- function(data, corrected_df){
    md <- summary(lmer(scale(dyad_perceived_min_t2) ~ scale(isc_recoded) + scale(dyad_perceived_min_t1) + same_class + same_middleschool + same_handedness + (1|subject1) + (1|subject2), data))

  B <- md$coefficients[[2,1]]
  #df <- md$coefficients[[2,3]]
  df <- corrected_df - 5
  p <- md$coefficients[[2,5]]
  t <- md$coefficients[[2,4]]
  p_penalized <- 2 * pt(-abs(t), df/2)
  
  return(data.frame(B, df, p, p_penalized))
}

results_isc_perceived_min <- ddply(master_df_doubled, .(brain_region), func_isc_perceived_min_z, corrected_df_n)
results_isc_perceived_min$corrected_p <- p.adjust(results_isc_perceived_min$p_penalized, method = "fdr")
results_isc_perceived_min_contr <- ddply(master_df_doubled, .(brain_region), func_isc_perceived_min_z_contr, corrected_df_n)
results_isc_perceived_min_contr$corrected_p <- p.adjust(results_isc_perceived_min_contr$p_penalized, method = "fdr")

results_isc_perceived_min
results_isc_perceived_min_contr
```

